04_limma_gon_F

### 2019-04-03 analysis/04_limma_gon_F.Rmd
female gonad without control

---
title: "04_limma_gon"
output: md_document
---

DESeq2 is _not_ recommended for experiments with more than 100 samples ([see Mike Love's post](https://mikelove.wordpress.com/2016/09/28/deseq2-or-edger/)), so I decided to try the limma package. I followed [this tutorial](https://www.bioconductor.org/packages/devel/workflows/vignettes/RNAseq123/inst/doc/limmaWorkflow.html).

@@

knitr::opts_chunk$set(fig.path = '../figures/gon/',cache=TRUE)

First, I read in the data I processed in 00_datawrangling.Rmd.

@@

```{r filter}
colData <- colData %>%
  dplyr::filter(grepl('gonad', tissue)) %>%
  droplevels()
row.names(colData) <- colData$V1

# print sample sizes
colData %>% select(sex,treatment, tissue)  %>%  summary()

@@

#  perform likelihood ratio test and thresholded testing
fit <- glmFit( parentalobject, parentaldesign, robust=T)
tr <- glmTreat(fit, lfc = 1)
topTags(tr)
```

@@

```{r plotMDS}
plotMDS(parentalobject, cex = 0.5)
```

@@

```{r plotMDS-lables}
levels(colData$treatment)
col.treatment <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6")[colData$treatment]

plotMDS(parentalobject,col=col.treatment, labels = colData$sex)
legend("bottom",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
title("Gonad Colored by Treatment")

plotMDS(parentalobject,dim=c(3,4), col=col.treatment, labels = colData$sex)
legend("topright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
title("Gonad Colored by Treatment")
```

@@

```{r 01-contrasts}
# view all levels
levels(colData$group)

# subset of conrasts - sex specific comparing hatch to lay
my.contrasts <- makeContrasts(
	         FG_CB = female.gonad.control - female.gonad.bldg,
	         FG_BL = female.gonad.bldg - female.gonad.lay,
	         FG_Li3 = female.gonad.lay - female.gonad.inc.d3,
	         FG_i39 = female.gonad.inc.d3 - female.gonad.inc.d9,
	         FG_i917 = female.gonad.inc.d9 - female.gonad.inc.d17,
	         FG_i17H = female.gonad.inc.d17 - female.gonad.hatch,
	         FG_H5 = female.gonad.hatch -  female.gonad.n5,
	         FG_n59 = female.gonad.n5 - female.gonad.n9,
	         FG_n9C = female.gonad.n9 - female.gonad.control,

	         MG_CB = male.gonad.control - male.gonad.bldg,
	         MG_BL = male.gonad.bldg - male.gonad.lay,
	         MG_Li3 = male.gonad.lay - male.gonad.inc.d3,
	         MG_i39 = male.gonad.inc.d3 - male.gonad.inc.d9,
	         MG_i917 = male.gonad.inc.d9 - male.gonad.inc.d17,
	         MG_i17H = male.gonad.inc.d17 - male.gonad.hatch,
	         MG_H5 = male.gonad.hatch -  male.gonad.n5,
	         MG_n59 = male.gonad.n5 - male.gonad.n9,
	         MG_n9C = male.gonad.n9 - male.gonad.control,

	         FG_n9B = female.gonad.n9 - female.gonad.bldg,
	         MG_n9B = male.gonad.n9 - male.gonad.bldg,
levels=parentaldesign)

@@

mycontrasts <- c("FG_CB", "FG_BL", "FG_Li3", "FG_i39", "FG_i917", "FG_i17H", "FG_H5", "FG_n59", "FG_n9C",
                 "MG_CB", "MG_BL", "MG_Li3", "MG_i39", "MG_i917", "MG_i17H", "MG_H5", "MG_n59", "MG_n9C",
                 "FG_n9B", "MG_n9B")

@@

# volcano plots

```{r volcanoplots}
# from http://www.compbio.dundee.ac.uk/user/pschofield/Projects/teaching_pg/workshops/biocDGE.html#maplots

lrt <- glmLRT(fit,coef=2)
topTags(lrt)

tt <- topTags(lrt,n=10000)$table

ggplot(data=tt) + geom_point(aes(x=logFC,y=-log(FDR),color=logCPM)) +
  scale_colour_gradientn(colours=c("#000000" ,"#FF0000" ))
```

>>>>

---
title: "04_limma_gon_F"
output: md_document
---

This anlaysis will _exclude_ the control timepoint and will example males and females separately.

@@

knitr::opts_chunk$set(fig.path = '../figures/gon_F/',cache=TRUE)
```

# Females first

@@

```{r filter}
colData <- colData %>%
  dplyr::filter(grepl('gonad', tissue)) %>%
  dplyr::filter(treatment != "control") %>%
  dplyr::filter(sex == "female") %>%
  droplevels()
row.names(colData) <- colData$V1

# print sample sizes
colData %>% select(treatment, tissue)  %>%  summary()

@@

#  perform likelihood ratio test and thresholded testing
fit <- glmFit( parentalobject, parentaldesign, robust=T)
tr <- glmTreat(fit, lfc = 1)
topTags(tr)
head(tr$table)
```

@@

```{r plotMDS}
levels(colData$treatment)
col.treatment <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6")[colData$treatment]
myfill <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6")

plotMDS(parentalobject, cex = 0.5, labels = colData$treatment, col=col.treatment)
title("Female Gonad MDS 1v2")
plotMDS(parentalobject, cex = 0.5, labels = colData$treatment, col=col.treatment, dim=c(3,4))
title("Female Gonad MDS 3v4")
plotMDS(parentalobject, cex = 0.5, labels = colData$treatment, col=col.treatment, dim=c(5,6))
title("Female Gonad MDS 5v6")
```

@@

@@

```{r MDSplots}
# view all levels
levels(colData$group)

# subset of conrasts - sex specific comparing hatch to lay
my.contrasts <- makeContrasts(
	         FG_BL = female.gonad.bldg - female.gonad.lay,
	         FG_Li3 = female.gonad.lay - female.gonad.inc.d3,
	         FG_i3i9 = female.gonad.inc.d3 - female.gonad.inc.d9,
	         FG_i9i17 = female.gonad.inc.d9 - female.gonad.inc.d17,
	         FG_i17H = female.gonad.inc.d17 - female.gonad.hatch,
	         FG_Hn5 = female.gonad.hatch -  female.gonad.n5,
	         FG_n5n9 = female.gonad.n5 - female.gonad.n9,

	         FG_Bi3 = female.gonad.bldg - female.gonad.inc.d3,
	         FG_Bi9 = female.gonad.bldg - female.gonad.inc.d9,
	         FG_Bi17 = female.gonad.bldg - female.gonad.inc.d17,
	         FG_BH = female.gonad.bldg - female.gonad.hatch,
	         FG_Bn5 = female.gonad.bldg - female.gonad.n5,
	         FG_Bn9 = female.gonad.bldg - female.gonad.n9,

	         FG_Li9 = female.gonad.lay - female.gonad.inc.d9,
           FG_Li17 = female.gonad.lay - female.gonad.inc.d17, 
	         FG_LH = female.gonad.lay - female.gonad.hatch,
	         FG_Ln5 = female.gonad.lay - female.gonad.n5,
	         FG_Ln9 = female.gonad.lay - female.gonad.n9,

	         FG_i3i17 = female.gonad.inc.d3 - female.gonad.inc.d17,
	         FG_i3H = female.gonad.inc.d3 - female.gonad.hatch,
	         FG_i3n5 = female.gonad.inc.d3 - female.gonad.n5,
	         FG_i3n9 = female.gonad.inc.d3 - female.gonad.n9,

	         FG_i9H = female.gonad.inc.d9 - female.gonad.hatch,
	         FG_i9n5 = female.gonad.inc.d9 - female.gonad.n5,
	         FG_i9n9 = female.gonad.inc.d9 - female.gonad.n9,

	         FG_i17n5 = female.gonad.inc.d17 - female.gonad.n5,
	         FG_i17n9 = female.gonad.inc.d17 - female.gonad.n9,

	         FG_Hn9 = female.gonad.hatch -  female.gonad.n9,

levels=parentaldesign)

colnames(my.contrasts)

@@

mycontrasts <- colnames(my.contrasts)
mycontrasts

@@

```{r volcanoplots}
for(i in mycontrasts){
  plotVolcanos(i)
}
```
